\begin{abstract}
We present a technique for automatic verification of concurrent
data structure algorithms that operate on different forms
of dynamically allocated heap structures such as singly-linked lists,
skiplists, and arrays of lists.
Such concurrent algorithms are notorciously difficult verify automatically,
since they often employ fine-grained synchronization, and since 
they consist of an unbounded number of concurrent threads, which
operate on an unbounded domain of data values, and use
unbounded dynamically allocated memory. 
All existing automated verification techniques
restrict attention to algorithms where
the heap represents shared data byp singly-linked
lists, often imposing additional restrictions.
In this paper, we present the first technique which can automatically verify concurrent
data structure algoritms that operate on dynamically allocated
heap structures which are more complex than just singly-linked lists.
Our approach is the first framework that
can automatically verify concurrent data structure algorithms that employ
skip lists, singly linked lists, as well as arrays of singly linked lists,
Our technique is based on a novel shape abstraction,
called {\em fragment abstraction} for representing
different kinds of unbounded heap structures.
%% such as arrays of singly linked lists and skip lists.
%% , for representing an unbounded heap.
%% a novel {\em shape abstraction}, which is designed to cope with an unbounded
%% heap (note that the heap visible to a single thread is still unbounded).
Fragment abstration represents  heap states by a set of {\em fragments},
each of which is a pair of node types (called tags) that are
connected by a pointer. A tag can be seen as a finitary abstraction of
a heap node, which summarizes both local information about
values of its data fields
%% (applying suitable data abstraction),
as well as global information about its position in the heap.
a set of fragments represents the set of heap structures in
which each pair of pointer-connected nodes are represented by some
fragment in the set.
Fragment abstraction can be combined with other abstractions for
handling unbounded data domains and for handling an unbounded number of threads.
We have implemented our framework in a tool and successfully used it to  prove linearizability  for a wide range of concurrent data structure algorithms.
In particular, we provide the first automated verification of such
algorithms that operate on skiplists and arrays of lists with timestamps.
\end{abstract}
