\paragraph{Related Work}

Much previous work has been devoted to
the {\it manual} verification of linearizability for
concurrent data structure implementations. Examples
include~\cite{LF:pldi13,Aaron:locigcal:linearizability}.
%% %
%% For instance, Turon {\it et al.} \cite{Aaron:locigcal:linearizability}
%% propose a semantic model, 
%% based on Kripke logical relations, that can be used for proving linearizability.
%% %
%% Liang {\it et al.} \cite{LF:pldi13}
%% propose a program logic with a lightweight instrumentation mechanism 
%% for verifying linearizability for algorithms with non-fixed LPs.
%
%% In \cite{OHearnlist}, O'Hearn {\it et al.}  define a
%% {\it hindsight lemma} that
%% provides a non-constructive evidence for linearizability. 
%% %
%% The lemma is used to prove linearizability of an optimistic variant of 
%% the lazy set algorithm.
%% by identifying the hindsight property of the algorithm. 
Vafeiadis \cite{Vafeiadis:Thesis}
uses forward and backward simulation relations together
with history or prophecy variables to prove linearizability.
%
These approaches are manual, and without
tool implementations.
{\it Mechanical} proofs of linearizability, using interactive theorem
provers, have been reported in 
\cite{Colvin:Lazy-List,Derrick:fm14,SWD:cav12,SDW:tcl14}.
%
For instance, Colvin {\it et al.} \cite{Colvin:Lazy-List}
verify the lazy set algorithm in  PVS,
using a combination of forward and backward simulations.


There are several works presenting techniques for {\it automated} verification
of functional correctness (in the sense of, e.g., linearizability)
for concurrent data structures.
All these techniques have been applied only to programs operating on singly linked
lists~\cite{YaSa:queue,Sagiv:correlation,Amit:comparisonAbstraction,EQSST:tacas10,Vafeiadis:cav10,AHHR:integrated,meyer:vmcai16,Quy:sas16}, often under restrictions, such as bounding the number of accessing
threads~\cite{Amit:comparisonAbstraction,Vechev:spin09,CernyRZCA:CAV10}.

\v{C}ern{\'y} {\it et al} \cite{CernyRZCA:CAV10} show decidability of a class
of concurrent data strcuture implementations with a bounded number of threads operating on a singly-linked list structure.

Our technique for specifying and verifying linearizability uses the
technique of observers~\cite{AHHR:integrated}, which adapts ideas from
the automata-theoretic approach~\cite{VW:modelchecking} and
data-independence~\cite{Wolper:dataindependence}. This technique
reduces the problem of checking linearizability to that of control-state
reachability.
We use further development of this approach, in particular that linearizability
of concurrent stacks and queues can be specified without requiring any user
annotation~\cite{BEEH:icalp15,HSV:concur13}.
To verify the linearizability of concurrent sets, we use
the light-weight instrumentation technique, extending the technique with
fixed linearization points, of~\cite{Quy:sas16}.

For sequential programs,
there are several techniques for automated verification of programs that operate
on heap structures that are more general than singly-linked lists.
These include approaches based on forest automata~\cite{forester12,boxes13,Quy:atva13:journal}, separation logic~ \cite{InvaderCAV08,dudka13},
automata \cite{artmc12}, or graph grammars \cite{juggrnaut10}, in some
cases also combined with data
abstraction~\cite{Loginov:AbstrRefViaInductLearning:05}.
These approaches have not been applied to verify functional properties of
concurrent programs operating on heap structures beyond singly-linked lists.

%% develops an automatic tool for proving  linearizability that
%% employs instrumentation to verify logically pure executions.
%% %
%% However, this work can handle non-fixed LPs only for read-only methods,
%% i.e, methods that do not modify the heap.
%% %
%% This means that the method cannot handle
%%  algorithms like the {\it Elimination} queue ~\cite{Shavit:ElimQueue}, {\it HSY} stack~\cite{HSYstack}, {\it CCAS}~\cite{Harris:CAS}, 
%% {\it RDCSS}~\cite{Harris:CAS} and {\it HM} set~\cite{ArtOfMpP} that we consider in this paper. In addition, their shape abstraction is not powerful enough to handle algorithms like {\it Harris} set~\cite{Harris:list} and {\it Michael} set~\cite{Michael:list} that are also handled by our method.
%% %
%% %For instance, in the {\it HSY} stack algorithm both
%% %the {\tt push} and {\tt pop} operations modify
%% %the heap and help each other~\cite{HSYstack}. 
%% %
%% Chakraborty {\it et al.} \cite{HSV:concur13}
%% describe an ``aspect-oriented'' method for  modular verification 
%% of concurrent queues that they use to prove linearizability of the Herlihy/Wing queue.
%% %
%% We can incorporate this technique into
%% our framework by a suitable construction of observers.
%% The method can not be applied to sets.
%% %
%% The most recent work of Zhu {\it et al.} \cite{Poling}
%% describe a tool that is applied for specific set, queue, and stack  
%% algorithms. For queue algorithms, their technique can handle queues with helping mechanism except for {\it HW} queue~\cite{HeWi:linearizability} which is handled by our paper.
%% %
%% For set algorithms, the authors can only handle those that perform an optimistic contains (or lookup) operation by applying the {\it hindsight lemma} from 
%% \cite{OHearnlist}. 
%% %
%% Hindsight-based proofs provide only {\it non-constructive} 
%% evidence of linearizability.
%% %
%% Furthermore, some algorithms (e.g., the unordered list algorithm
%% considered in Sec.~\ref{section:experiments} of this paper)
%% do not contain the code patterns required by the hindsight method. Algorithms with non-optimistic contains (or lookup) operation like {\it HM}~\cite{ArtOfMpP}, {\it Harris}~\cite{Harris:list} and {\it Michael}~\cite{Michael:list} sets cannot be verified by their technique.  
%% %
%% Vechev {\it et al.}~\cite{Vechev:spin09}
%% check linearizability with user-specified non-fixed LPs,
%% using a tool for finite-state verification.
%% Their method assumes a bounded number of threads, and
%% they report state space explosion when having more than two threads.
%% %
%% Dragoi  {\it et al.} \cite{Henzinger:CAV13} describe a method for proving
%% linearizability that is applicable to algorithms with non-fixed LPs.
%% %
%% However, their method needs to rewrite the implementation so that all operations 
%% have linearization points within the rewritten code.
%% %
%% \v{C}ern{\'y} {\it et al} \cite{CernyRZCA:CAV10} show decidability of a class
%% of programs with a bounded number of threads operating on concurrent data structures.
%% %
%% %% This paper is about a different problem
%% %% Horn and Kroening \cite{Kroening-Linearizability:FORTE15}
%% %% have recently presented an efficient linearizability checker based on {\it P-compositionality} 
%% %% which can be applied when having a bounded number of threads.
%% %
%% Finally, the works~\cite{AHHR:integrated,BLMRS:cav08,Vafeiadis:vmcai09}
%% all require fixed linearization points.

We have not found any report in the literature of a
verification method that is sufficiently powerful to
automatically verify properties such as linearizablity
for concurrent data structure implementations based on heap structures, which
are more general than singly-linked lists. In particular, no approaches
for automatically verifying concurrent skiplist-based algorithms
~\cite{Fomitchev:2004,ArtOfMpP,Linden:opodis13,Sundell:2005} or arrays of lists operating on time stamps~\cite{ts-stack}.
%reported in
%Section~\ref{section:experiments} (as also confirmed by the CAVE
%tool implementors).
%The same applies to
%the tool in~\cite{Poling}.



